%!TEX TS-program = xelatex
\documentclass{nihbiosketch}

%------------------------------------------------------------------------------

\name{Brokamp, Cole}
\eracommons{brokampr}
\position{Assistant Professor}

\begin{document}
%------------------------------------------------------------------------------

\begin{education}
University of Cincinnati  & B.S           & 06/2010  & Biomedical Engineering \\
University of Cincinnati               & Ph.D.         & 04/2016  & Biostatistics and Bioinformatics \\
Cincinnati Children's Hospital Medical Center  & Postdoctoral Research Fellow  & 10/2017  & Biostatistics and Epidemiology \\
\end{education}

\section{Personal Statement}

\begin{statement}

I am an environmental and population health scientist trained at the
intersection of biostatistics, epidemiology, and geoinformatics in order to
advance precision public health. My work has focused on using
GIS, machine learning, and causal inference methods with ``big spatial data'' to
estimate the effect of spatiotemporal environmental exposures on neurobehavioral and
psychiatric pediatric health
outcomes. Specifically, my interest lies in the
interaction of these urban exposures with individual- and community-level
socioeconomic characteristics in order to to identify subpopulations of children
who are more susceptible to environmental health problems and inform targeted
interventions.

In this project, I will serve as a co-Investigator and will utilize my
expertise and experience in clincial machine learning prediction models to
develop and implement an early warning system for units at
CCHMC. Working with the other investigators, I will contribute to Aim 2 by
developing time-series and machine learning models to identify possible
interventions to reduce injury rates and also to predict injury rates specific
to each unit and to the hospital overall.  I will also work to integrate these
models into an interactive heatmap dashboard and early notifications that will
comprise the early-warning system developed in Aim 3. My experience in
developing temporal and spatiotemporal machine learning prediction models will
allow me to successfully collaborate as a part of this team and contribute to
successful completion of the study aims.\\

  \begin{enumerate}

	\item \textbf{Cole Brokamp}, Roman Jandarov, Monir Hossain, Patrick Ryan. Predicting Daily Urban Fine Particulate Matter Concentrations Using Random Forest. Environmental Science \& Technology. 52 (7). 4173-4179. 2018.

	\item \textbf{Cole Brokamp}, MB Rao, Patrick Ryan, Roman Jandarov. A comparison of resampling and recursive partitioning methods in random forest for estimating the asymptotic variance using the infinitesimal jackknife. \textit{Stat}. 6(1). 360-372. 2017.

  \item \textbf{Cole Brokamp}, Jeffrey R. Strawn, Andrew F. Beck, Pat Ryan.
    Pediatric Psychiatric Emergency Department Utilization and Fine
    Particulate Matter: A Case-Crossover Study. \textit{Environmental Health
      Perspectives}. 2019.
    
  \end{enumerate}

\end{statement}

%------------------------------------------------------------------------------
\section{Positions and Honors}

\subsection*{Positions and Employment}
\begin{datetbl}
2012--2016 & Research Associate, Department of Environmental Health, University of Cincinnati \\	
2016--2017  & Research Fellow, Cincinnati Children's Hospital Medical Center Division of Biostatistics \& Epidemiology\\
2017--      & Assistant Professor of Pediatrics, the University of Cincinnati Department of Pediatrics and Cincinnati Children’s Hospital Medical Center Division of Biostatistics \& Epidemiology\\
\end{datetbl}

\subsection*{Honors}
\begin{datetbl}
2010            & B.S. awarded with Distinguished Honors, University of Cincinnati\\
2016            & CCHMC Division of Biostatistics \& Epidemiology Travel Award\\
2016            & CCHMC Arnold W. Strauss Fellowship Award\\
2017            & CCHMC Epidemiology \& Biostatistics Top Publication and Top
Research Achievement\\
\end{datetbl}

%------------------------------------------------------------------------------

\section{Contribution to Science}

\begin{enumerate}

\item My early career has been spent developing exposure assessment models for
  spatialtemporal pollutants and community characteristics based on machine learning
  techniques.  This work includes the first machine
  learning or ensemble model used to assess exposure to elemental components of
  particulate matter. Recent introduction of remote sensing satellite data has
  allowed for extension of the land use random forest model to produce daily
  estimates of air pollution back to 2000 at a resolution of 1 x 1 km.

\begin{enumerate}
	
  \item \textbf{Cole Brokamp}, Eric B. Brandt, Patrick H. Ryan. Assessing
  Exposure to Outdoor Air Pollution for Epidemiological Studies:
  Model-based and Personal Sampling Strategies. \emph{Journal of Allergy
    and Clinical Immunology}. 2019.

	\item \textbf{Cole Brokamp}, Roman Jandarov, Monir Hossain, Patrick Ryan. Predicting Daily Urban Fine Particulate Matter Concentrations Using Random Forest. Environmental Science \& Technology. 52 (7). 4173-4179. 2018.

	\item \textbf{Cole Brokamp}, Roman Jandarov, MB Rao, Grace LeMasters, Patrick Ryan. Exposure assessment models for elemental components of particulate matter in an urban environment: A comparison of regression and random forest approaches. \textit{Atmospheric Environment}. 151. 1-11. 2017.
	
	\item \textbf{Cole Brokamp}, MB Rao, Patrick Ryan, Roman Jandarov. A comparison of resampling and recursive partitioning methods in random forest for estimating the asymptotic variance using the infinitesimal jackknife. \textit{Stat}. 6(1). 360-372. 2017.

  \item \textbf{Cole Brokamp}, Grace LeMasters, Patrick Ryan. Residential
    mobility impacts exposure assessment and community socioeconomic
    characteristics in longitudinal epidemiology studies. \emph{Journal of
      Exposure Science and Environmental Epidemiology}. 26(4). 428-34. 2016.

  \item \textbf{Cole Brokamp}, MB Rao, Tina Zhihua Fan, Patrick H Ryan. Does the
    elemental composition of indoor and outdoor PM2.5 accurately represent
    the elemental composition of personal PM2.5?. \emph{Atmospheric
      Environment}. 101. 226-234. 2015.

\end{enumerate}

\item Building on advanced exposure assessment models has allowed me to lead epidemiological studies on the impacts of the built environment (e.g., fine particulate matter, greenspace, combined sewer overflows, community deprivation) on several different pediatric health outcomes (e.g., psychiatric, neurobehavioral, gastrointestinal, and all-cause hospital utilization).

  \begin{enumerate}

  \item \textbf{Cole Brokamp}, Jeffrey R. Strawn, Andrew F. Beck, Pat Ryan.
    Pediatric Psychiatric Emergency Department Utilization and Fine
    Particulate Matter: A Case-Crossover Study. \textit{Environmental Health
      Perspectives}. 2019.

  \item Juliana Madzia, Patrick Ryan, Kimberly Yolton, Zana Percy, Nick Newman, Grace
    LeMasters, \textbf{Cole Brokamp}. Residential Greenspace Is Associated with Childhood
    Behavioral Outcomes. \textit{Journal of Pediatrics}. 30. 37-43. 2019.

  \item \textbf{Cole Brokamp}, Andrew F. Beck, Neera K. Goyal, Patrick Ryan,
    James M. Greenberg, Eric S. Hall. Material Community Deprivation and
    Hospital Utilization During the First Year of Life: An Urban
    Population-Based Cohort Study. \textit{Annals of Epidemiology}. 30. 37-43.
    2019.

  \item \textbf{Cole Brokamp}, Andrew F. Beck, Louis Muglia, Patrick Ryan. Combined Sewer Overflow Events and Childhood Emergency Department Visits: A Case-Crossover Study. \textit{Science of the Total Environment}. 607-608. 1180-1187. 2017.
    
  \item Kim Hartley, Patrick Ryan, \textbf{Cole Brokamp}, Gordon L. Gillespie.
    Effect of greenness on asthma in children: A systematic review.
    \emph{Public Health Nursing}. 0. 1-8. 2020.

    \item Rebecca Gernes, \textbf{Cole Brokamp}, Glenn Rice, J. Michael Wright,
    Michelle Kondo, Yvonne Michael, Geoffrey Donovan, Demetrios Gatziolis,
    David Bernstein, Grace LeMasters, James Lockey, G. Khurana Hershey,
    Patrick Ryan. Using high-resolution residential greenspace measures in
    an urban environment to assess risks of allergy outcomes in a cohort of
    children. \emph{Science of the Total Environment}. 668. 760-767. 2019.

  \item Kelly J Brunst, Patrick H Ryan, \textbf{Cole Brokamp}, David Bernstein,
    Tiina Reponen, James Lockey, Gurjit K Khurana Hershey, Linda Levin,
    Sergey A Grinshpun, Grace LeMasters. Timing and duration of
    traffic-related air pollution exposure and the risk for childhood wheeze
    and asthma. \emph{American Journal of Respiratory and Critical Care
      Medicine}. 192(4). 421-427. 2015.

  \end{enumerate}

\item I have developed a novel approach and accompanying software package called DeGAUSS
  which overcomes the multiple privacy-related challenges in the use of address data in
  multi-site studies and also serves as a more general reproducible research
  tool for geocoding and geomarker assessment. This approach is currently being
  implemented in a wide variety of national environmental health studies. Extending this
  approach into a scalable and sustainable framework for automated integration of
  disparate and heterogeneous geomarkers via spatiotemporal location has reduced
  the need for manual data curation and specialized expertise required
  to utilize them within biomedical research studies.

\begin{enumerate}
	
	\item \textbf{Cole Brokamp}, Chris Wolfe, Todd Lingren, John Harley, Patrick Ryan. Decentralized and Reproducible Geocoding and Characterization of Community and Environmental Exposures for Multi-Site Studies. \textit{Journal of American Medical Informatics Association.} 25(3). 309-314. 2018.
	
	\item \textbf{Cole Brokamp}. DeGAUSS: Decentralized Geomarker Assessment for Multi-Site Studies. \textit{Journal of Open Source Software}. 2018. 

\end{enumerate}

\item I have also contributed to several studies on the disparities of health outcomes within children and the contribution of the place-based and social determinants of health to these disparities in order to identify root causes and meaningful solutions.

\begin{enumerate}

  \item Sarah Orkin, \textbf{Cole Brokamp}, Toshifumi Yodoshi, Andrew T. Trout,
  Chunyan Liu, Syeda Meryum, Stuart Taylor, Christopher Wolfe, Rachel
  Sheridan, Aradhna Seth, Mohammad Alfrad Nobel Bhuiyan, Sanita Ley, Ana
  Catalina Arce-Clachar, Kristin Bramlage, Robert Kahn, Stavra Xanthakos,
  Andrew F. Beck, Marialena Mouzaki. Community Socioeconomic Deprivation
  and Non-Alcoholic Fatty Liver Disease Severity. \emph{Journal of
    Pediatric Gastroenterology and Nutrition}. 2019.

  \item Erica Andrist, \textbf{Cole Brokamp}, Stuart Taylor, Carley Riley,
  Andrew Beck. Neighborhood Poverty and Pediatric Intensive Care Use.
  \textit{Pediatrics}. 2019.

	\item Andrew F. Beck, Carley L. Riley, Stuart Taylor, \textbf{Cole Brokamp},
    Robert S. Kahn. Toward a Culture of Health in Hospitals: Pervasive
    population disparities in inpatient bed-day rates across conditions and
    subspecialties. \textit{Health Affairs}. 37(4). 551-559. 2018.
		
	\item Lauren C. Riney, \textbf{Cole Brokamp}, Andrew F. Beck, Wendy Pomerantz,
    Hamilton Schwartz, Todd A. Florin. Emergency Medical Services Utilization is
    Associated with Community Deprivation in Children. \textit{Prehospital
      Emergency Care}. 2018.

\end{enumerate}

\item Lastly, I have contributed to a research team that has recently used
  functional data analysis combined with joint modeling (FD-JM) to identify and
  predict rapid decline in lung function among patients with cystic fibrosis
  (CF) lung disease. My work in translating this predictive model into an interactive application has allowed for patients and clinicians to take advantage of it at the bedside.  Focus groups and partnerships with clinicians have allowed us to iteratively develop the application based on end-user feedback. Work with the CF Foundation Patient Registry (CFFPR) to implement these models and visualizations into clinical settings has improved prognostic care.

\begin{enumerate}
	
  \item Rhonda D. Szczesniak, Weiji Su, \textbf{Cole Brokamp}, Ruth H. Keogh,
  John P. Pestian, Michael Seid, Peter J. Diggle, John P. Clancy. Dynamic
  predictive probabilities to monitor rapid cystic fibrosis disease
  progression. \emph{Statistics in Medicine}. 2019.

	\item Rhonda D. Szczesniak, \textbf{Cole Brokamp}, Weiji Su, Gary L. McPhail,
    John Pestian, and John P. Clancy. Improving Detection of Rapid Cystic
    Fibrosis Disease Progression—Early Translation of a Predictive Algorithm
    into a Point-of-Care Tool. \textit{IEEE Journal of Translational Engineering
      in Health and Medicine.} 7(1). 1-8. 2019.

	\item Rhonda Szczesniak, \textbf{Cole Brokamp}, Weiji Su, Gary L. McPhail, John Pestian, John P. Clancy. Early Detection of Rapid Cystic Fibrosis Disease Progression Tailored to Point of Care: A Proof-of-Principle Study. \textit{Healthcare Innovations and Point of Care Technologies}. (HI-POCT), 2017 IEEE. 204-207. 2017.

	\item Rhonda D. Szczesniak, Dan Li, Weiji Su, \textbf{Cole Brokamp}, John Pestian, Michael Seid, John P. Clancy. Phenotypes of Rapid Cystic Fibrosis Lung Disease Progression during Adolescence and Young Adulthood. \textit{American Journal of Respiratory And Critical Care Medicine}. 196(4). 471-478. 2017.
	
\end{enumerate}

\end{enumerate}

\subsection*{Complete List of Published Work in ORCiD:} 
\url{https://orcid.org/0000-0002-0289-3151}

\section{Additional Information: Research Support and/or Scholastic Performance}

\subsection*{Ongoing Research Support}

\bigskip

\textbf{ECHO Opportunities and Infrastructure Fund Award}\\
\emph{Decentralized and Reproducible Geomarker Assessment for Multi-Site
  Studies}\\
Brokamp, PI (09/01/2019 - 08/31/2021)\\
This award will work towards building geospatial exposure assessment
computing tools for utilizing high resolution spatiotemporal gridded
datasets within ECHO.\\
Role: PI

\bigskip


\textbf{ODH Contract No. CSP907820}\\
\emph{Model Identifying Geographic Areas in Ohio for Blood Lead Testing}\\
Brokamp, PI (04/01/2020 - 09/30/2020)\\
This award will develop a predictive model to determine which children should be
tested for potentially high blood lead during physician visits based on their
residential location. \\
Role: PI

\bigskip

\textbf{NIH 5UG3OD023282-02}\\
\emph{Children's Respiratory Research and Environment Workgroup
(CREW)}\\
Gern, PI (09/01/2016 - 08/31/2023)\\
This consortium will leverage a large (nearly 9000 births and
long-term follow-up of 6000-7000 children and young adults) and diverse
national data set, to harmonize data related to asthma clinical
indicators and early life environmental exposures, develop
standardized measures for prospective data collection, and conduct targeted enrollment of
additional subjects into existing cohorts.\\
Role: Co-I

\bigskip

\textbf{NIH/NHLBI R01HL141286-01A1}\\
\emph{Mapping Environmental Contributions to Rapid Lung Disease
Progression in Cystic Fibrosis}\\
Sczcesniak, PI (01/01/2019 - 12/31/2023)\\
The overall objective of this research is to leverage a rich CF
registry, extant national and local environmental data sources, and
prospectively collected study data to accurately forecast the onset of
rapid decline progression.\\
Role: Co-I

\bigskip

\textbf{NIH/NIEHS R21ES030092}\\
\emph{Developing and Evaluating Novel Strategies for Reporting Back Individual
  Results of Personal Air Monitors}\\
Ryan, PI (09/01/2019 - 08/31/2021)\\
The objectives of this proposal are to collaborate with study participants, caregivers and community stakeholders to develop effect report-back strategies for personal air pollution monitoring results and evaluate report-back materials and their influence on knowledge, behavior, and personal UFP exposure.\\
Role: Co-I

\bigskip

\textbf{NIH/NIA R21AG057983}\\
\emph{A Novel Research Infrastructure Enabling Life-Course Studies of
Healthy Aging}\\
Woo/Urbina, PI (08/15/2018 - 07/31/2023)\\
The goal of this two-phase study is to develop the data and biospecimen
infrastructure for the Bogalusa Heart Study, the Princeton Lipid
Research Study and the NHLBI Growth and Health Study and to
conduct pilot evaluations of the feasibility, acceptability and validity
of data collected using a variety of biometric sensors relating to
cardiometabolic risk, sleep quality and cognition.\\
Role: Co-I

\bigskip

\subsection*{Recently Completed Research Support}

\bigskip

\textbf{Internal Processes and Methods Award - Center for Clinical \&
  Translational Science \& Training}\\
\emph{Using Machine Learning to Supplement Electronic Health Record
  databases with Individual Socioeconomic Status}\\
Brokamp, PI (09/1/2017 - 06/30/2019)\\
Retrospective epidemiological studies are often created using electronic
health record databases that lack individual level demographic data. We
propose a novel machine learning based approach that uses open city and
auditor databases to predict individual level income and family
socioeconomic status.\\
Role: PI

\bigskip

\textbf{Internal Arnold W. Strauss Fellowship Award - Cincinnati
Children's Hospital}\\
\emph{Assessing Exposure to Air Pollution Across Time and Space}\\
Brokamp, PI (07/01/16 - 06/30/17)\\
The primary objective of this award is to combine satellite-based
measurements, land use characteristics, and meteorologic data to create
a hybrid spatiotemporal model for ground level exposure to particulate
matter using exact addresses and dates.\\
Role: PI

\bigskip

\textbf{Internal Processes and Methods Award - Center for Clinical \&
Translational Science \& Training}\\
\emph{Validating a Geocoding Approach for Multi Site Studies}\\
Brokamp, PI (01/24/17 - 06/30/17)\\
The primary objective of this award is to compare the geocoding
(assigning latitude and longitude coordinates to addresses) accuracy of
our software DeGAUSS (DEcentralized Geomarker Assessment for mUlti Site
Studies) to with other common geocoding software. Furthermore, each
method will be evaluated based on it ability to correctly estimate
environmental exposures and community-level characteristics.\\
Role: PI

\bigskip

\end{document}